---
title: "Fitness Data Prediction"
author: "Sammie Bae"
date: "December 22, 2015"
output: html_document
---

####By Sammie Bae
#Background:
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, data from belt accelerometers to predict what excercise 6 participants are doing between forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

#Goal
The goal of the project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. These exercise vary accordingly:  "exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E). Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes." The model should predict accurately on 20 sample different test cases.

#Setup
Loading the data and setup:
```{r}
library(knitr)
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(rattle)

#set the working directory
setwd("C:/Users/Sam.E/Desktop/datasciencecoursera/Practical Machine Learning/Course Write-up")
#Data Source:
if(!file.exists("pml-training.csv")){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")
}
if(!file.exists("pml-testing.csv")){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
}
training <-read.csv("pml-training.csv", na.strings=c("NA",""),stringsAsFactors = FALSE)
testing <-read.csv("pml-testing.csv", na.strings=c("NA",""),stringsAsFactors = FALSE)
dim(training);dim(testing)
```

#Data Preprocessing & Clean-up
```{r}
#undefined data to N/A
  undefinedToNA <- sapply(training, function(x) x=="#DIV/0!")
  training[undefinedToNA] <- NA
```  

To build a machine learning model, numeric values are preffered,
binary values such as Yes or No should be changed to 1 or 0
```{r}
#convert yes/no into 1/0
  #testing
    testing$new_window = 1*(testing$new_window=="yes")
    testing$new_window <- as.factor(testing$new_window)
  #training
    training$new_window = 1*(training$new_window=="yes")
    training$new_window <- as.factor(training$new_window)
    training$classe <- factor(training$classe)
```

There are lot of columns with N/A values that will interfere with our machine learning model.
These columns and the irelevant ID column should be removed from both the training and the testing data set.
```{r}
#removes first column with ID and username
training<- training[,-2]
testing <- testing[,-2]
#remove N/A value columns
training <- training[, unlist(lapply(training, function(x) !any(is.na(x))))]
testing <- testing[, unlist(lapply(testing, function(x) !any(is.na(x))))]
dim(training);dim(testing)
```

#Data Partitioning
For this partitioning, training and testing data set are 60% and 40% respectively
```{r}
inTrain<-createDataPartition(y=training$classe, p=0.6, list=FALSE)
dataTraining<-training[inTrain, ]
dataTesting <-training[-inTrain, ]
```

```{r}
training<- training[, sapply(training, is.numeric)]
testing <- testing[, sapply(testing, is.numeric)]
dim(dataTraining);dim(dataTesting)
```

#Model Building
##Method 1: Random Forrests
Random forest trees are one of the best prediction model in machine learning. This dataset will first be modelled with trees:
```{r}
trainingModelRF <- train(classe ~ .,trControl=trainControl(method = "cv", number = 4),data=dataTraining, method="rf")
trainingModelRF

```


```{r}
modelFitTraining <- train(classe ~ ., method="rpart", data=dataTraining)
fancyRpartPlot(modelFitTraining$finalModel)
```


```{r}
treeModel <- rpart(classe ~ ., data=dataTraining, method="class")
prp(treeModel)
```

```{r}
controlDTree <- trainControl(method="cv", 5)
modelDTree <- train(classe ~ ., data=dataTraining, method="rf", trControl=controlDTree, ntree=150)
modelDTree
```

```{r}
predictionRF<-predict(trainingModelRF, dataTesting)
confusionMatrix(predictionRF, dataTesting$classe)
accuracy <- postResample(predictionRF, dataTesting$classe)
model_accuracy <- accuracy[[1]]*100
model_accuracy
```
#Cross Validation

#Errors

#Prediction
Final prediction model will be applied to 20 data sets
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
#pml_write_files(predictionFinal)
```
#Summary