---
title: "Reproducible Research Assignment 1"
author: "Sammie Bae"
date: "April 23, 2016"
output: html_document
---

1. Load the data & Process/transform the data

Download the data from url if it doesn't exist in directory
```{r}
  setwd("C:\\Users\\Sam.E\\Desktop\\datasciencecoursera\\Reproducible Research\\Assignment 1")
  getwd()
  fileName <- "getdata_dataset.zip"
```

Download
```{r}
  if (!file.exists(fileName)){
    url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    f <- file.path(getwd(), "getdata_dataset.zip")
    download.file(url, f)
  }
```

Unzip it
```{r}
  if (!file.exists("repdata-data-activity")) { 
    unzip(fileName) 
  }
```

read the CSV data
```{r}
  repData <- read.csv("activity.csv")
```

2. What is mean total number of steps taken per day?
```{r}
  library(ggplot2)
  step.mean     <- tapply(repData$steps, repData$date, FUN=sum, na.rm=TRUE)
  aggregatedDay <- aggregate(steps ~ date, repData, as.vector)
  
  qplot(step.mean, binwidth=1000, xlab="Steps by day")
  
  mean(step.mean, na.rm=TRUE)
  median(step.mean, na.rm=TRUE)
```
Answer: the mean is approximately  9354 steps per day and the median is 10395 steps per day


3. What is the average daily activity pattern?
```{r}
  library(ggplot2)
  averageInterval<-aggregate(x=list(steps=repData$steps), by=list(interval=repData$interval), FUN=mean, na.rm=TRUE)
  aesOption <- aes(x=interval, y=steps)
  ggplot(data=averageInterval, aesOption) + geom_line()+xlab("5-minute Intervals")+ylab("Average Steps Taken")
```

4. Calculate Imputing missing values
```{r}
  missing <- is.na(repData$steps)
  table(missing)
  
```
Answer: there are 2304 missing values that need to be filled:
```{r}
  fillValues <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
      filled <- c(steps)
    else
      filled <- (averageInterval[averageInterval$interval==interval, "steps"])
    return(filled)
  }
  filledData <- repData
  filledData$steps <- mapply(fillValues, filledData$steps, filledData$interval)
  
  total.steps <- tapply(filledData$steps, filledData$date, FUN=sum)
  qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
  mean(total.steps)
  median(total.steps)
```
  
5. Are there differences in activity patterns between weekdays and weekends?
```{r}
  weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    weekends <-c("Saturday", "Sunday")
    weekdays <-c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
    if (day %in% weekends)
      return("weekend")
    else if (day %in% weekdays)
      return("weekday")
    else
      stop("invalid date")
  }
  filledData$date <- as.Date(filledData$date)
  filledData$day <- sapply(filledData$date, FUN=weekday.or.weekend)
  
  par(mfrow = c(2, 2))
   
  averages <- aggregate(steps ~ interval + day, data=filledData, mean)
  ggplot(data=averageInterval, aesOption) + geom_line()+xlab("5-minute Intervals")+ylab("Average Steps Taken")
  ggplot(averages, aes(interval, steps)) + geom_line() + xlab("5-minute interval") + ylab("Number of steps")
  
```

Answer: yes, there is generally more activity on weekends