---
title: "Motor Trend Data Analysis"
output: html_document
---
####By Sammie Bae
###Executive Summary:
In this analysis, looking at a data set of a collection of cars, relationship between a set of variables and miles per gallon (MPG) (outcome) will be explored to answer the following questions:
```{}
1. “Is an automatic or manual transmission better for MPG”
2. "Quantify the MPG difference between automatic and manual transmissions"
```

###Data Set:
```{r}
mtcars
```

###1. “Is an automatic or manual transmission better for MPG”
am in the data column represents automatic transmission. With cars of am 1 being manual and cars of am 0 being automatic


```{r}
autoCars  <- subset(mtcars,am==0) 
manualCars<- subset(mtcars,am==1)
```

manualCars is non-automatic cars and manualCars is automatic cars


```{r}
boxplot(autoCars$mpg,manualCars$mpg, las = 1, ylab ="Miles Per Gallon", xlab ="Transmission Type (Figure 1)", names=c("Automatic","Manual")) 
```



```{r}
summary(autoCars$mpg)
summary(manualCars$mpg)
```
by performing basic analysis, it is observed that min, median, mean, and max of manual cars are all greater than that of automatic cars as shown in Figure 1. This strongly supports that manual cars are more fuel efficient than automatic ones

####t.test verification
```{r}
t.test(mtcars$mpg~mtcars$am,var.equal=T)
```
To verify, t.test was conducted on the dataset .
The confidence interval (95%) does not contain zero (-10.84837, -3.64151) and p-value is greater then 0.005. Therefore, it can concluded that the average consumption of manual cars is lower than that of automatic cars.

###2. "Quantify the MPG difference between automatic and manual"
###Model Selection:

###Model 1: Simple Linear Model
```{r}
modelSimple <- lm(mpg~am, data = mtcars)
summary(modelSimple)
```

```{}
Adjusted R-squared:  0.3385
```
Based on this linear model, coefficient of determination only explains 33.8% of the variance.


###Model 2: Mutivariable Linear regression on all variables:
```{r}
mtcarsMultivarRegresion <- lm(mpg ~ ., data = mtcars)
summary(mtcarsMultivarRegresion)
```

From this summary, it can be concluded that weight (wt) is the factor which has the most correlation with mpg as it has the highest absolute coefficient. Other two significant factors are automatic transmission (am) and qsec (1/4 mile time).

However, there may be overfitting from including all variables. Therefore, more precise model must be formed:

```{}
step function selects a formula-based model by AIC
```

```{r}
modelStep = step(lm(data = mtcars, mpg ~ .),trace=0,steps=99999)
summary(modelStep)
```


```{}
Adjusted R-squared:  0.8336
```
This model shows again that weight (wt), automatic transmission (am), and qsec (1/4 mile time) are the most significant factors for mpg.
In addition, this model now has coefficient of determination that explains 83.36% of the variance.

Using this information, final model can be formed:
```{r}
modelFinal <- lm(mpg~am + wt + qsec, data = mtcars)
summary(modelFinal)

```

###Conclusion
Therefore, the model suggests that manual transmissions have a better mpg than automatic transmissions by an average of 2.9 mpg. However, weight has even more significant impact at 3.9165 mpg and 1/4 mile time also has considerable impact at 1.2259 mpg.




